---
title: "ASSIGNMENT 4"
author: "Luke Syverson"
date: '2023-04-23'
output:
  word_document: default
  pdf_document: default
  html_document: default
bibliography: "F:/OneDrive/Documents/School/Github Desktop/dsc520/assignments/assignment04/bibliography.bib"
---

# Markdown Basics

## Favorite Foods
  1) Beef Steak
  2) Fish
  3) Sourdough Bread
## Images
![](~/School/Github Desktop/dsc520/completed/assignment04/plots)
All Cases (Log Plot)
## Add a Quote
> “Out of all the things I have lost, I miss my mind the most.” -- Mark Twain

## Add an Equation
$$\frac{1+ \sqrt5}{2}$$
## Add a Footnote

This is a footnote[^1].

[^1]:This is the above footnote description.

## Add Citations

* R for Everyone (@lander2014r)
* Discovering Statistics Using R (@field2012discovering)

# Inline Code
```{r include=FALSE}
setwd("~/School/Github Desktop/dsc520")
heights_df <- read.csv("data/r4ds/heights.csv")
covid_df <- read.csv("data/nytimes/covid-19-data/us-states.csv")
covid_df$date <- as.Date(covid_df$date)
california_df <- covid_df[ which( covid_df$state == "California"), ]
ny_df <- covid_df[ which(covid_df$state == "New York"),]
florida_df <- covid_df[ which(covid_df$state == "Florida"),]
library(ggplot2)
```
## NY Times COVID-19 Data
```{r echo=FALSE}
ggplot(data=florida_df, aes(x=date, group=1)) +
  geom_line(aes(y = cases, colour = "Florida")) +
  geom_line(data=ny_df, aes(y = cases,colour="New York")) +
  geom_line(data=california_df, aes(y = cases, colour="California")) +
  scale_colour_manual("",
                      breaks = c('Florida', 'New York', 'California'),
                      values = c('Florida' = 'darkred', 'New York' = 'darkgreen', 'California' = 'steelblue')) +
  xlab(" ") + ylab("Cases") + scale_y_log10()
```
## R4DS Height vs Earnings
```{r echo=FALSE}
ggplot(heights_df, aes(x=height, y=earn, col=sex)) + geom_point() + xlab("Height") + ylab("Earnings") + ggtitle('Height vs. Earnings')
```
# Tables
```{r}
name <- c("Aragon", "Bilbo", "Frodo", "Galadriel", "Sam", "Gandalf", "Legolas", "Sauron", "Gollum")
race <- c("Men", "Hobbit", "Hobbit", "Elf", "Hobbit", "Maia", "Elf", "Maia", "Hobbit")
in_fellowship <- c(TRUE, FALSE, TRUE, FALSE, TRUE, TRUE, TRUE, FALSE, FALSE)
ring_bearer <- c(FALSE, TRUE, TRUE, FALSE, TRUE, TRUE, FALSE, TRUE, TRUE)
age <- c(88, 129, 51, 7000, 36, 2019, 2931, 7052, 589)
characters_df <- data.frame(name, race, in_fellowship, ring_bearer, age)
```
## Knitr Table with Kable
```{r}
library(stringr)
knitr::kable(characters_df, caption = 'One Ring to Rule Them All'
             , col.names = str_to_title(gsub("_", " ", names(characters_df))))
```
## Pandoc Table
```{r}
library(pander)
pandoc.table(characters_df, style = 'rmarkdown')
# I'm not sure if this is the right way to do this, but it works!
```
# References